---
title: "STAT448 - Assignment 1"
author: "Dipin P Joseph - 72746678, Nandakumar T - 35044765"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Question 1
#### c)
#### Defining matrix of Predictor Variable, X
```{r}
X <- matrix(c(3,4,5),nrow = 3)
X
```
#### Defining matrix of Outcome Variable, Y
```{r}
Y <- matrix(c(5,9,13),nrow = 3)
Y
```
#### Defining X matrix with an extra column of 1's
```{r}
one_col <- rep(1,3)
X_1 <- matrix(rbind(one_col,c(3,4,5)),nrow = 3,byrow = TRUE)
X_1
```
#### Calculating Beta matrix with the equation - B=((X'X)^-1)(X'Y)
```{r}
B<-solve(t(X_1)%*%(X_1))%*%t(X_1)%*%Y
B
```
#### Coefficients for linear regression, Slope-B1, Intercept-B0
```{r}
B0 = B[1]
B1 = B[2]
B0
B1
```
#### Finding residuals of the linear regression model, Res = Y-(B1*X+B0) which equals zero.
```{r}
Y_Pred <-(B1*X+B0)
Res = Y-Y_Pred
Res
```
### Question 1
#### d)
#### Performing linear regression with function, lm() and displaying it's summary
```{r}
l_reg <- lm(formula = Y ~ X)
summary(l_reg)
```
#### From the summary, R square value is 1 and residual error 0 which implies that model is a perfect fit but highly dependent on perdictor variable and is unrealistic.
#### Plotting X,Y, predicted values and linear model.
```{r echo=FALSE}
# Scatter plot for X and Y
plot(main = 'Linear Regression Y~X',X,Y, xlim = c(0,6) , ylim = c(-10,15), pch=19, col="#13871c", cex=1)
# Predicted values for X with regression
Y_pred <- predict.lm(l_reg, newdata = list(x=X))
# Points with X and predicated vaues of Y
points(X, Y_pred, col="#d37024", cex=2)
# Linear regression line
abline(l_reg, col="#3f4da7", lwd=3)
legend(0,15, legend=c("Predicted Y", "Actual Y"),
       col=c("#d37024", "#13871c"), pch=c(1,19), cex=0.8)
text(x=1, y=-2, labels="y = β0 + β1X = -7+4x", col="#3f4da7",srt=27)
text(x=0.55, y=-2.75, labels="^     ^      ^", col="#3f4da7",srt=27)
```

### Question 2
#### b)
#### Defining matrix of Predictor Variable, X
```{r}
X <- matrix(c(2,2,2),nrow = 3)
X
```
#### Defining matrix of Outcome Variable, Y
```{r}
Y <- matrix(c(5,9,13),nrow = 3)
Y
```
#### Defining X matrix with an extra column of 1's
```{r}
one_col <- rep(1,3)
X_1 <- matrix(rbind(one_col,c(2,2,2)),nrow = 3,byrow = TRUE)
X_1
```
#### Calculating Beta matrix with the equation - B=((X'X)^-1)(X'Y)
```{r}
tryCatch({
  B<-solve(t(X_1)%*%(X_1))%*%t(X_1)%*%Y
}, error = function(error_condition) {
  print(paste("Error:  ",error_condition))
})
det(t(X_1)%*%X_1)
``` 
#### Since det(X_1) being a zero(singular matrix), inverse of (X_1'X_1) is not defined. Hence, slope of regression line can't be determined.
#### Performing linear regression with function, lm() and displaying it's summary
```{r}
l_reg <- lm(formula = Y ~ X)
summary(l_reg)
```
#### The linear regression is failing and slope(b1) can't be found due to singularity issue. Residual error for the model is 4 on df=2.
#### Plotting X,Y, predicted values and linear model,
```{r echo=FALSE}
# Scatter plot for X and Y
plot(main = 'Linear Regression Y~X',X,Y, xlim = c(0,4) , ylim = c(-5,15), pch=19, col="#13871c", cex=1)
# Predicted values for X with regression
Y_pred <- predict.lm(l_reg, newdata = list(x=X))
# Points with X and predicated vaues of Y
points(X, Y_pred, col="#d37024", cex=2)
legend(0,15, legend=c("Predicted Y", "Actual Y"),
       col=c("#d37024", "#13871c"), pch=c(1,19), cex=0.8)
text(x=2.5, y=10, labels="y = β0 + β1X", col="#3f4da7")
text(x=2.45, y=10.8, labels="^     ^      ^", col="#3f4da7")
```

#### Predictions from this rank-deficient fit may be misleading. 
